Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-27 docs/paper.tex
Finished Section 4

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{bruin2021} with various parameters.

@@

	The first experiments were performed on a 2.2GHz Intel i7-8750H hexacore with 16GiB of memory. For each setting, a fixed subset of 50 out of 200 instances was used. %Hyperthreading was used in experiments, but because agent-settings were solved in sequence as batches, it is expected that this has affected experimental outcomes equally.

	The second experiments comparing the different MAPFM algorithms within the peer group were run on a server with a virtualized Intel Xeon E5-2683 dodecacore operating at 2.0GHz with 8GiB of RAM. All 200 instances were used for each setting.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with three teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of 75\% maps with one team that exhaustive methods can not within the time-out (\ref{fig:i-75-1-p}).

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that ordered enumeration works well in practice. Simple pruning with child pruning outperforms enhanced pruning but this needs to be qualified. Comparing Exh+S+B+O with Exh+E+B+O shows that part of the difference can be explained by the use of simple pruning instead of enhanced pruning. Nonetheless, in instances with 75\% wall (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}) there appears to be a real benefit to searching a smaller ICT. Across ICTS-m variants, similar trends can be observed but the differences across pruning strategies are too slight to draw any conclusions from. As discussed in Section \ref{future}, more work is needed to compare simple child pruning with other pruning methods for ICTS.

	The benefit of using ID together with ordered enumeration is not clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause of the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, it is unlikely that there will be a significant benefit to using ID, as agents from the same team planned independently often route to the same goal.

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching \cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching \cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching \cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams \cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is because from each family of algorithms one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. The initial groups are here the teams, which is why in single team scenarios (Figures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}) matching ID has no effect. Hypothetically, matching ID could explain why both A*+ID+OD and EPEA* outperform the methods using just sorted exhaustive matching in maps with 75\% wall and three teams (\ref{fig:r-75-3-p}).

	CBM solving all problems with one team (Figure \ref{fig:r-25-1-p} and \ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving scenarios with a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

>>>>

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI maps \cite{sturtevant2012} are often used and potentially a new standard benchmark for MAPFM could be defined based on these maps, similar to an existing standard benchmark for MAPF \cite{stern2019}; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was already a challenge computationally. This observation informed the creation of a map generator by the author of \cite{bruin2021} with various parameters.

@@

	The first experiment was performed on a 2.2GHz Intel i7-8750H hexacore with 16GiB of memory. For each setting, a fixed subset of 50 out of 200 instances was used. %Hyperthreading was used in experiments, but because agent-settings were solved in sequence as batches, it is expected that this has affected experimental outcomes equally.

	The second experiment comparing different MAPFM algorithms was run on a server with a virtualized Intel Xeon E5-2683 dodecacore operating at 2.0GHz with 8GiB of RAM. All 200 instances were used for each setting.

@@

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	To investigate the relative performance of various algorithms discussed in Section \ref{section:icts-matching}, a selection of both exhaustive ICTS and ICTS-m variants was made. Their performance in terms of fraction of problems solved within a timeout of 120s is shown in Figure \ref{fig:i-probs}. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (Figure \ref{fig:i-probs}), with performance being more similar in the maps with 75\% wall. In particular on 75\% wall maps with three teams (Figure \ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases; additionally, some ICTS-m variants manage to solve a small fraction of 75\% maps with one team that exhaustive methods can not within the time-out (Figure \ref{fig:i-75-1-p}). This can be explained by the complexity of ICTS-m not being directly related to the number of matchings in contrast to the complexity of exhaustive ICTS.

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by Figure \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that ordered enumeration works well in practice. Simple pruning with child pruning outperforms enhanced pruning but this needs to be qualified. Comparing Exh+S+B+O with Exh+E+B+O shows that part of the difference can be explained by the use of simple pruning instead of enhanced pruning. Nonetheless, in instances with 75\% wall (Figures \ref{fig:i-75-1-p},\ref{fig:i-75-3-p}) there appears to be a real benefit to searching a smaller ICT. Across ICTS-m variants, similar trends can be observed but the differences across pruning strategies are too slight to draw any conclusions from. As discussed in Section \ref{future}, more experiments are needed to adequately compare simple child pruning with other pruning methods for ICTS.

	The benefit of using ID together with ordered enumeration is not clear. Ordered enumeration with ID performs much better in a setting with more obstacles and more teams (Figure \ref{fig:i-75-3-p}), but ordered enumeration without ID on the other hand is superior in an open setting with a single team (Figure \ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause of the slowdown observed once $M$ grows large (Figures \ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as between ICTS-m and ICTS-m+ID. Another factor is that if all agents belong to the same team, the benefit of using ID is reduced as agents from the same team planned independently often route to the same goal.

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching \cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching \cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching \cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams \cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is because from each family of algorithms one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. The initial groups are here the teams, which is why in single team scenarios (Figures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}) matching ID has no effect. Hypothetically, matching ID could explain why both A*+ID+OD and EPEA* outperform the methods using just sorted exhaustive matching in maps with 75\% wall and three teams (Figure \ref{fig:r-75-3-p}).

	CBM solving all problems with one team (Figures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving scenarios with a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.
